\section{Gridcoin Grants and Resource Allocation}

\label{sec:grants}

A blockchain based financial system which prioritizes the pragmatic and democratized advancement of data based research holds the potential to revolutionize the funding process for computational science, data analysis, and simulation based research.

\subsection{A Touch of General Research and Grant History}

If history is our teacher, it can be said that basic research holds high value with regard to scientific endeavors and progress.  Basic research is the general exploration of a subject without the quest for a final product.  It is led by the innate inquisitive drive that encourages many of humanity’s actions.  In general, basic research leads to discoveries and creations which have connections to a wide array of subjects other than that which was intended to be explored.  In other words, learning a little more about X can lead us to discover Y, Z, A, B, and C.  For example, the basic research of mathematics in the 19th and early 20th century is the reason we have computers today.  The research into symmetry will give us quantum computings and entirely new realms of physics.  The research into the stickleback fish gave us a large portion of the evolutionary biology knowledge base which is helping to cure disease today.

It is difficult to exactly pinpoint when, but over time basic research has become less of a focus when compared to translational research.  Translational research, also known as applied research, is research which can be quickly translated into a marketable product.  The main effect of this transition has been a shift of funding sources from the public sector to the private sector.  This in turn has had many detrimental side-effects.  For example, universities are now allowed to establish and run their own for-profit research entities.  In the past, they could function only at the whim of taxpayer funding, which meant the research belonged to citizens.  In the past, private corporations were barred from influencing public university researchers.  Now, it would be a challenge to find a public university that does not host corporations which require professors and researchers to sign Non Disclosure Agreements.

The world of research funding has changed over the past decades.  Gridcoin seeks to change it again for the coming decades.

\subsection{The Current Grant System}

Research is an expensive endeavor that need not guarantee results.  Consider the largest research projects of the millenium, the Large Hadron Collider and ATLAS.  There is no guaranteed product, marketable result, or promise other than that of discovery.  Unfortunately, this mentality does not correlate into general grant based research.  

In order to receive funding, a researcher must submit a grant proposal to an entity with available funds.  In the past, these entities were governments.  In a democracy, this means the funding came from and the science belonged to the citizens of a country.  Over time, government funding was decreased and taken up by private for-profit entities.  As a result, the focus of research has shifted from basic, or foundational research, to translational, or applied research.  Foundational research gives us the knowledge required to build new things.  Applied research tries to build new things to be sold on a market.  Both are important, however foundational is required for applied to exist, and in general it is fairly difficult to monetize and own foundational research.

The shift from public grants to private grants means that most research is owned by private, for-profit entities.  The for-profit nature of the entities putting up funds for grants means that researchers are encouraged to focus on the rapid and hyper-focused creation of marketable products.  This structure also limits the funding available for a large swath of researchers who seek to perform foundational research as foundational research often does not result in quick to market products.

\subsection{The Gridcoin Grant System}

A system which allows and encourages researchers to request funding from a general population encourages the spread of knowledge and understanding.  This peer-to-peer funding structure has been tested and developed in the venture capital, banking and loan, and manufacturing industries.  Gridcoin seeks to bring it to computational science, data analysis, and simulation based research.

Gridcoin presents a blockchain and network which allocates funds and resources (IPP) to foundational and applied research based on the desires of the network.  These desires are influenced by individual preference, external factors, and the ability of a researcher or entity to educate network participants to the usefulness of their research.  For example, during times of dramatic shifts in climate, projects with a focus on weather modeling and climate prediction might see more funding and IPP directed toward their completion.  Of those projects, those best able to communicate their purpose and the benefits of their completion will likely receive the majority of these resources.

\subsection{Where Do Funds Come From?}

https://github.com/gridcoin-community/Gridcoin-Tasks/issues/191