\section{Idle Processing Potential}

\label{sec:IPP}

Any technologically advanced society produces a certain amount of unused processing power as a waste product of its hardware.  Gridcoin aims to use this idle processing power to run its network, maximizing the use of resources in already existing processors.  Idle Processing Potential, or IPP, can be measured as the unused processing power of a cpu over time.  It is difficult to quantify the Idle Processing Potential which exists in the world today, as research on the subject is sparse.  It is possible, however to get a feel of the enormous potential presented by an IPP driven network.  Every phone, every computer, every gaming system, every car, refrigerator, toy and object that houses a CPU is potentially part of this network. 
 
To explore the potential of an IPP driven network, let us try to find the approximate IPP of all smartphones in circulation today and compare the smartphone network’s IPP with BOINC and Bitcoin.  We will have to guess for some variables as the data is not readily available.

BOINC currently hosts 18 petaflops on its network with 171,457 active volunteers.  In comparison, the Bitcoin network hosts more than 80 exaflops.  This is about 100 times more powerful than the top 500 centralized supercomputers combined, which sit at a modest 748.4 petaflops as of June 2017.  An iPhone 6, on the other hand, offers up to 7 gigaflops.  There are approaching 2.5 billion smartphones in circulation.  Assuming 7 gigaflops as an average, the processing potential of all smartphones in circulation would then be 2.5 billion multiplied by 7 gigaflops, or 17.5 exaflops.  Let us say that smartphones are idle at least 50 percent of the time, when the user is asleep.  The IPP of the smartphone network would then be at least 8.75 exaflops.  Better data is clearly required to eliminate assumptions in these calculations and determine the true IPP of the smartphone network.  Directing this and the IPP currently wasted by other CPU enabled hardware toward scientific, social, and economic endeavors presents truly inconceivable possibilities in the realms of social understanding, science, and equality. 